---
title: "ds607_week7_JagdishChhabria"
author: "Jagdish Chhabria"
date: "March 16, 2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# This is the code chunk for reading in the HTML file into R as a data frame
library(XML)
library(RCurl)
library(rvest)
library(dplyr)

#fileurl<-"C:\\Jagdish\\Masters Programs\\CUNY\\Data 607 Data Acquisition and Management\\Week7\\books.html"
fileurl<-getURL("https://raw.githubusercontent.com/Jagdish16/jagdish_r_repo/master/DATA607/Week7/books.html")

#books1<-readHTMLTable(doc=fileurl, trim=T, as.data.frame=T, header=T, Encoding="windows-1252")
#books2<-htmlParse(file=fileurl, encoding = "windows-1252", as.data.frame=T)
#books2<-read_html(fileurl, encoding = "UTF-8")
#tables<-html_nodes(books2, "table")
# table1 <- html_table(tables, fill = TRUE)

booksdf <- as.data.frame(read_html(fileurl) %>% html_table(fill=TRUE))
booksdf
```




```{r}
# This is the code chunk for reading in the XML file into R as a data frame
library(XML)
library(RCurl)
library(httr)

#fileurl<-"C:\\Jagdish\\Masters Programs\\CUNY\\Data 607 Data Acquisition and Management\\Week7\\books1.xml"

fileurl<-getURL("https://raw.githubusercontent.com/Jagdish16/jagdish_r_repo/master/DATA607/Week7/books1.xml")

books<-xmlParse(file=fileurl)
books_df<-xmlToDataFrame(books, stringsAsFactors = FALSE)
books_df


```

```{r}
# This is the code chunk for reading in the JSON file into R as a data frame
library(jsonlite)
library(dplyr)
library(RCurl)

#fileurl<-"C:\\Jagdish\\Masters Programs\\CUNY\\Data 607 Data Acquisition and Management\\Week7\\books5.json"
fileurl<-getURL("https://raw.githubusercontent.com/Jagdish16/jagdish_r_repo/master/DATA607/Week7/books5.json")

books.df<-fromJSON(fileurl) %>% as.data.frame
books.df


```

The 3 dataframes are not completely identical. There are minor differences such as: 1) the "id" column missing in the dataframe converted from the html file; 2) capitalization of column names is different across the datafarmes created from html versus xml, while the column names for the dataframe created from the json file are pre-fixed by the name of the first object in the json file. Presumably these differences are on account of how the different handling functions work under the hood.




