---
title: 'DS606_Lab5_Inference for numerical data_JagdishChhabria'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## North Carolina births

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## Exploratory analysis

Load the `nc` data set into our workspace.

```{r load-data, eval=TRUE}
load("more/nc.RData")

```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | -----------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1.  What are the cases in this data set? How many cases are there in our sample?
The cases are individual births, and there are 1000 cases in the sample.


```{r}
nrow(nc)
```

As a first step in the analysis, we should consider summaries of the data. This
can be done using the `summary` command:

```{r summary, eval=TRUE}
summary(nc)
```

As you review the variable summaries, consider which variables are categorical 
and which are numerical. For numerical variables, are there outliers? If you 
aren't sure or want to take a closer look at the data, make a graph.

```{r}
str(nc)

```
The following variables are numerical: fage, mage, weeks, visits, gained, weight. Looking at the summary data, there seem to be some outliers for mother's age and father's age.



Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot 
highlight about the relationship between these two variables?


```{r}
boxplot(nc$weight~nc$habit)
```
The plot shows that the median weights are pretty close for both categories (smoker and non-smoker), though the median weight is slightly higher for non-smoking mothers. Both categories have some outliers, though there seem to be may more outliers for the non-smoking mothers.  


The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following function to 
split the `weight` variable into the `habit` groups, then take the mean of each
using the `mean` function.

```{r by-means, eval=TRUE}
by(nc$weight, nc$habit, mean)
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test
.

## Inference

3.  Check if the conditions necessary for inference are satisfied. Note that 
you will need to obtain sample sizes to check the conditions. You can compute 
the group size using the same `by` command above but replacing `mean` with 
`length`.

```{r}
by(nc$weight, nc$habit, length)
```
The sample size is adequate for making inferences. In both cases (smoker and non-smoker), it is above 30, and less than 10% of the population of mothers. It is mentioned that the sample was randomly drawn.

4.  Write the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different.

H0: avg.weight.smoker - avg.weight.nonsmoker = 0
Ha: avg.weight.smoker - avg.weight.nonsmoker <> 0


Next, we introduce a new function, `inference`, that we will use for conducting
hypothesis tests and constructing confidence intervals. 

```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```

Let's pause for a moment to go through the arguments of this custom function. 
The first argument is `y`, which is the response variable that we are 
interested in: `nc$weight`. The second argument is the explanatory variable, 
`x`, which is the variable that splits the data into two groups, smokers and 
non-smokers: `nc$habit`. The third argument, `est`, is the parameter we're 
interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next
we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a 
confidence interval (`"ci"`). When performing a hypothesis test, we also need 
to supply the `null` value, which in this case is `0`, since the null 
hypothesis sets the two population means equal to each other. The `alternative`
hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` 
of inference can be `"theoretical"` or `"simulation"` based.

5.  Change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to smoking and 
non-smoking mothers.
```{r}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```



By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$)
. We can easily change this order by using the `order` argument:

```{r inf-weight-habit-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```

* * *

## On your own

-   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.

```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = mean(nc$weeks), 
          alternative = "twosided", method = "theoretical" 
         )
```
From the above, it can be seen that the 95% confidence interval for the average length of the pregnancy in the population is between 38.15 and 38.51 weeks. 

-   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.
```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = mean(nc$weeks), 
          alternative = "twosided", method = "theoretical", conflevel=0.90
         )
```


-   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.

```{r}
mean(nc$gained, na.rm = TRUE)
nc.new<-subset(nc,is.na(nc$gained)==0, select=c(gained, mature))
by(nc.new$gained, nc.new$mature, mean)
```

H0: avg.gain.mature - avg.gain.younger = 0
Ha: avg.gain.mature - avg.gain.younger <> 0


-   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.

```{r}
nc.age<-subset(nc, is.na(nc$mage)==0,select=c(mage, mature))
by(nc.age$mage, nc.age$mature, max)
by(nc.age$mage, nc.age$mature, min)
```
Based on the above, it can be seen that the maximum age for younger mothers is 34 years, and the minimum age for mature mothers is 35 years. So the cutoff is 35 years.


-   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.

```{r}
boxplot(nc$weight~nc$premie)
```

H0: avg.weight.fullterm - avg.weight.premature = 0

Ha: avg.weight.fullterm - avg.weight.premature <> 0

The above hypotheses test whether the weight of the baby at birth is affected by whether it is full-term or pre-mature.


```{r}
inference(y = nc$weight, x = nc$premie, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")
```

```{r}
inference(y = nc$weight, x = nc$premie, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```
Based on both hypothesis test and confidence intervals, the null hypothesis can be rejected at the 95% confidence level. The null value of 0 is not contained in the confidence interval, and the p-value associated with the test-statistic is 0.





<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab 
written by the faculty and TAs of UCLA Statistics.
</div>
